---
title: "Assignment 3 Task 3: Santa Barbara Urchins "
author: "Alia Ajina"
date: "10/28/2021"
output: html_document
---

# Overview
This code uses version control, project management, data wrangling, analysis, and visualization skills to explore red and purple urchin records from the Santa Barbara Coastal Long Term Ecological Research (SBC LTER) program's divers. The data is from a long-term experiment on kelp removal and the sea urchin (red and purple) size frequency distribution.

Data Citation: Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b. 

# Setup

Attach tidyverse, here, janitor, and kableExtra packages.

```{r setup, include=TRUE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
library(lubridate)
```

# Section 1

## Read in Urchins data

```{r, Read in Data}
urchins <- read_csv(here("data", "urchins.csv"))
```

## Data Wrangling

Clean the data

```{r}
urchins_clean <- urchins %>% 
  clean_names()
```

## Data Visualization

Create a finalized, faceted histogram comparing the distribution of sizes between purple and red sea urchins.

```{r}
# Plot histogram and change formatting accordingly
ggplot(data = urchins_clean, aes(x = size)) +
  geom_histogram(aes(fill = common_name, color = common_name), bins = 10) + 
  facet_wrap(~ common_name) + # Facet wrap by "common_name"
  scale_color_manual(values = c("white", "white")) +
  scale_fill_manual(values = c("#CBC3E3", "#801B1B")) +
  theme_gray() + 
  theme(axis.title.x = element_text(hjust = 0.5, vjust = 0.5),
        plot.title = element_text(hjust = .5), 
        plot.subtitle = element_text(hjust = .5),
        plot.caption = element_text(hjust = .3),
        legend.position = "none") +
  labs( x = "Urchin Diameter (cm)",
        y = "Number of Urchins",
        title = "Distribution of Sizes between Red and Purple Urchins (AJINA)",
        subtitle = "Santa Barbara Coastal Long Term Ecological Research", 
        caption = "Figure 1: Distribution of sizes between purple and red sea urchins at five different sites 
        along the Santa Barbara Coast.")
```

## Summary Statistics

Summary table of statistics for purple and red urchins including mean, median, standard deviation, and sample size. Use KableExtra to finalize the table.

```{r}
# Create a summary statistics table

urchins_summary <- urchins_clean %>% 
  group_by(common_name) %>% 
  summarize(mean = round(mean(size), 2), 
            median = median(size),
            sd = round(sd(size), 1),
            sample_size = n())

# Use KableExtra to finalize the table

urchins_summary %>% 
  kable(col.names = c("Common Name", 
                      "Mean Size (cm)",
                      "Median Size (cm)",
                      "Standard Deviation (cm)",
                      "Sample Size"),
        caption = "<center>Table 1: Summary Statistics of Purple and Red Urchins<center>") %>% 
  kable_styling(bootstrap_options = "bordered",
                full_width = F, position = "center")


```
The summary statistics for red and purple sea urchins allows us to see that red sea urchins are slightly larger than the purple urchins, at 2.3 cm larger on average leading to a 41% difference in mean urchin diameter between the two. The purple sea urchins have a smaller standard deviation and a mean and median that are nearly the same. In comparison to red sea urchins, this concludes that purple sea urchin data has less spread than red sea urchin data.

# Section 2

## Data Wrangling

Create a new subset of data, purple_urchins, convert the date column to the class "Date", create a new column that only contains the year, and and create another new column called site_full that contains the full name of the five sites, not just their abbreviation.

```{r}
# Create new subset
purple_urchins <- urchins_clean %>% 
# Convert date column to class "Date"
  filter(common_name == "Purple Urchin") %>% 
  mutate(date = mdy(date)) %>%
  mutate(year = year(date)) %>% 
# Create new column with full site names
  mutate(site_full = case_when(site == "NAPL" ~ "Naples Reef", 
                               site == "AQUE" ~ "Arroyo Quemado Reef",
                               site == "CARP" ~ "Carpinteria Reef",
                               site == "MOHK" ~ "Mohawk Reef",
                               site == "IVEE" ~ "Isla Vista Reef"))
```

### Table of Purple Urchins observed by site and year.

```{r}
# Create table

total_purple_urchins <- purple_urchins %>% 
  group_by(site_full, year) %>% 
  summarize(count=n())
```

## Data Visualization

Two finalized graphs of counts of purple urchins recorded at each site over time for the years included in the study visualized in both a bar graph and a line graph for optimal comparison.

### Bar Graph of Purple Urchin Distribution per Year by Site

```{r}
# Create bar graph
ggplot(data = total_purple_urchins, 
       aes(x = year, y = count)) +
# Change aesthetics accordingly
  geom_col(aes(color = site_full, fill = site_full)) +
  facet_wrap(~ site_full, scale = 'free') + # Facet wrap by "site_full" and change scale to free in order to have an x and y axis on each individual plot
  scale_color_manual(values = c("white", "white", "white", "white", "white")) +
  scale_fill_manual(values = c("#B7E9F7", "#90EE90", "#F4F186", "#FED8B1", "#FFB6C1")) + # Change colors of faceted histograms and their outline for better visuals
  theme_gray() + 
  scale_x_continuous(limits=c(2007,2020), 
                     breaks = c(2008, 2010, 2012, 2014, 2016, 2018, 2020)) + 
  scale_y_continuous(limits=c(0,1200)) + # Add limits to the scales in order to keep the axis the same
  labs( x = "Year",
        y = "Number of Purple Urchins",
        title = " Distribution of Purple Urchins from 2008 - 2020 by site (AJINA)",
        caption = "Figure 1.1: Distribution of purple sea urchins in each site from the years 2008 to 2020 from 
        SBC LTER data") +
# Adjust positioning
  theme(axis.text.x = element_text(size = 6, hjust = 0.6, angle = 50),
        axis.text.y = element_text(size = 6),
        axis.title.x = element_text(hjust = 0.5, 
                                    vjust = 0.5),
        plot.title = element_text(hjust = .5), 
        plot.subtitle = element_text(hjust = .5),
        plot.caption = element_text(hjust = .2), 
        legend.position = "none")
```

### Line graph of Purple Urchin Distribution per Year by Site

```{r}
# Create line graph
ggplot(data = total_purple_urchins, 
       aes(x = year, y = count)) + 
  geom_line(aes(color = site_full)) + # Adjust aesthetics and visuals according to preferences
  scale_color_manual(values = c("#B7E9F7", "#90EE90", "#F4F186", "#FED8B1", "#FFB6C1")) + # Change colors of line
  theme_minimal() +  # add breaks to the scale in order to have uniform x and y axis
  scale_x_continuous(breaks = c(2008, 2010, 2012, 2014, 2016, 2018, 2020)) +
  scale_y_continuous(breaks = c(0, 250, 500, 750, 1000, 1250, 1500)) +
  labs( x = "Year",
        y = "Number of Purple Urchins",
        title = " Distribution of Purple Urchins from 2008 - 2020 by site (AJINA)",
        caption = "Figure 1.1: Distribution of purple sea urchins in each site from the years 
        2008 to 2020 from SBC LTER data",
        color = "Site Location",
        Key = "Site") +
# Adjust placement/orientation of text 
  theme(axis.text.x = element_text(size = 6, hjust = 0.6),
        axis.text.y = element_text(size = 6),
        axis.title.x = element_text(hjust = 0.5, 
                                    vjust = 0.5),
        plot.title = element_text(hjust = .5), 
        plot.subtitle = element_text(hjust = .5),
        plot.caption = element_text(hjust = .3))
```

Two graphs, a faceted histogram and a line graph, were created for optimal comparison. The graphs reflect a decrease in the urchin population since 2008. However, the intense drop in 2020 may likely be due to a lack in data. Data stops on February 11th, near the start of Covid-19 in which the project was likely shut down for a period of time. As such the data may not be reflective of actual conditions. Another factor to consider is that the treatment on each site was ignored. We did not separate the controls, where kelp was not removed, and the annuals where kelp was removed. To determine accurate representations of the changes in counts of purple sea urchins, we would need to separate these two treatments. Overall decreases in purple sea urchin populations could also be attributed to sampling error or other environmental factors.
